Probabilistic Forecasts in Hierarchical~Time~Series

### 2018-04-17 ProbabilisticCoherence.tex
Changes to Gaussian example, drop subscripts, notational changes, work on qf separately

\def\PQ{\begin{pmatrix}\bm{P}\\[-0.2cm]\bm{Q}\end{pmatrix}}
\def\bt{\begin{pmatrix}\tilde{\bm{b}}_{t+h}\\[-0.2cm]\tilde{\bm{t}}_{t+h}\end{pmatrix}}

@@

Suppose $\mathcal{N}(\hat{\bm{\mu}}_{t+h}, \hat{\bm{\Sigma}}_{t+h}) \overset{d}{\leftrightarrow} \hat{\bm{f}}(\hat{\bm{y}}_{t+h})$ is an incoherent forecast distribution at time $t+h$. Then from \eqref{4.5} it follows that
\begin{equation}
\bm{f_B}(\tilde{\bm{b}}_{t+h},\tilde{\bm{t}}_{t+h})=\hat{\bm{f}}(\bm{S}\tilde{\bm{b}}_{t+h}+\bm{R}\tilde{\bm{t}}_{t+h}) \, \Big|\bm{S} ~ \vdots~ \bm{R}\Big| = \frac{\hat{\bm{f}}(\bm{S}\tilde{\bm{b}}_{t+h}+\bm{R}\tilde{\bm{t}}_{t+h}) }{\Big|(\bm{S} ~ \vdots~ \bm{R})^{-1}\Big|}.
\end{equation}
By substituting the Gaussian distribution function for $\bm{f_B}(\cdot)$ we get
\begin{align*}
\bm{f_B}(\cdot)
& =
\frac{\exp\left\{-\frac{1}{2}(\bm{S}\tilde{\bm{b}}_{t+h}+\bm{R}\tilde{\bm{t}}_{t+h}-\bm{\hat{\mu}}_{t+h})' \bm{\hat{\Sigma}_{t+h}}^{-1}(\bm{S}\tilde{\bm{b}}_{t+h}+\bm{R}\tilde{\bm{t}}_{t+h}-\bm{\hat{\mu}}_{t+h})\right\}}{(2\pi)^{\frac{n}{2}}\Big|\bm{\hat{\Sigma}}_{t+h}\Big|^{\frac{1}{2}}\Big|(\bm{S} ~ \vdots~ \bm{R})^{-1}\Big|},\\
& =
\frac{\exp\left\{-\frac{1}{2}\Big((\bm{S} ~ \vdots~ \bm{R})\bt-\bm{\hat{\mu}}_{t+h}\Big)' \bm{\hat{\Sigma}_{t+h}}^{-1}\Big((\bm{S} ~ \vdots~ \bm{R})\bt-\bm{\hat{\mu}}_{t+h}\Big)\right\}}{(2\pi)^{\frac{n}{2}}\Big|\bm{\hat{\Sigma}}_{t+h}\Big|^{\frac{1}{2}}\Big|(\bm{S} ~ \vdots~ \bm{R})^{-1}\Big|},\\
& =
\frac{1}{(2\pi)^{\frac{n}{2}}\Big|\bm{\hat{\Sigma}}_{t+h}\Big|^{\frac{1}{2}}\Big|(\bm{S} ~ \vdots~ \bm{R})^{-1} \Big|}
\exp \Big\{-\frac{1}{2}\Big(\bt-(\bm{S} ~ \vdots~ \bm{R})^{-1}\bm{\hat{\mu}}_{t+h}\Big)'\\
& \hspace*{6cm} \Big[(\bm{S} ~ \vdots~ \bm{R})\bm{\hat{\Sigma}_{t+h}}(\bm{S} ~ \vdots~ \bm{R})'\Big]^{-1}
\Big(\bt-(\bm{S} ~ \vdots~ \bm{R})^{-1}\bm{\hat{\mu}}_{t+h}\Big) \Big\}.
\end{align*}
Recall that
$$
(\bm{S} ~ \vdots~ \bm{R})^{-1} =
\begin{pmatrix}(\bm{R}'_\bot \bm{S})^{-1}\bm{R}'_\bot \\ \cdots \\ (\bm{S}'_\bot \bm{R})^{-1}\bm{S}'_\bot \end{pmatrix} =
\begin{pmatrix}
\bm{P} \\\bm{Q}
\end{pmatrix},
$$
where $\bm{P}=(\bm{R}'_\bot \bm{S})^{-1}\bm{R}'_\bot$ and $\bm{Q}=(\bm{S}'_\bot \bm{R})^{-1}\bm{S}'_\bot$. Then
\begin{align*}
\bm{f_B}(\cdot)& =\frac{1}{(2\pi)^{\frac{n}{2}}\Big|\bm{\hat{\Sigma}}_{t+h}\Big|^{\frac{1}{2}}\Big|\PQ \Big|}
\exp \Big\{-\frac{1}{2}\Big[\bt-\PQ\bm{\hat{\mu}}_{t+h}\Big]'\\[-0.5cm]
& \hspace*{7cm}
\Big[\PQ\bm{\hat{\Sigma}_{t+h}}\PQ'\Big]^{-1}\Big[\bt-\PQ\bm{\hat{\mu}}_{t+h}\Big] \Big\},\\[0.5cm]
 & =\frac{1}{(2\pi)^{\frac{n}{2}}\Big|\PQ\bm{\hat{\Sigma}_{t+h}}\PQ'\Big|^{\frac{1}{2}}}
\exp \Big\{-\frac{1}{2} \begin{pmatrix}\tilde{\bm{b}}_{t+h} - \bm{P}\bm{\hat{\mu}}_{t+h} \tilde{\bm{t}}_{t+h}- \bm{Q}\bm{\hat{\mu}}_{t+h}\end{pmatrix}' &\\[-0.5cm]
& \hspace*{7cm}
 \Big[\PQ\bm{\hat{\Sigma}_{t+h}}\PQ'\Big]^{-1}\begin{pmatrix}\tilde{\bm{b}}_{t+h} - \bm{P}\bm{\hat{\mu}}_{t+h}\\ \tilde{\bm{t}}_{t+h}- \bm{Q}\bm{\hat{\mu}}_{t+h}\end{pmatrix} \Big\}.
\end{align*}
Since $\Big[\PQ\bm{\hat{\Sigma}_{t+h}}\PQ'\Big] = \begin{pmatrix}
\bm{P}\bm{\hat{\Sigma}_{t+h}}\bm{P}' & \bm{P}\bm{\hat{\Sigma}_{t+h}}\bm{Q}' \\
\bm{Q}\bm{\hat{\Sigma}_{t+h}}\bm{P}' & \bm{Q}\bm{\hat{\Sigma}_{t+h}}\bm{Q}'
\end{pmatrix}$ we have
\begin{align*}
\bm{f_B}(\cdot)&=\frac{1}{(2\pi)^{\frac{n}{2}}\Big|\begin{pmatrix}
  \bm{P}\bm{\hat{\Sigma}_{t+h}}\bm{P}' & \bm{P}\bm{\hat{\Sigma}_{t+h}}\bm{Q}' \\
  \bm{Q}\bm{\hat{\Sigma}_{t+h}}\bm{P}' & \bm{Q}\bm{\hat{\Sigma}_{t+h}}\bm{Q}'
  \end{pmatrix}\Big|^{\frac{1}{2}}}
\exp \Big\{-\frac{1}{2} \begin{pmatrix}\tilde{\bm{b}}_{t+h} - \bm{P}\bm{\hat{\mu}}_{t+h}\\ \tilde{\bm{t}}_{t+h}- \bm{Q}\bm{\hat{\mu}}_{t+h}\end{pmatrix}'\\
&\hspace*{7cm}\begin{pmatrix}
\bm{P}\bm{\hat{\Sigma}_{t+h}}\bm{P}' & \bm{P}\bm{\hat{\Sigma}_{t+h}}\bm{Q}' \\
\bm{Q}\bm{\hat{\Sigma}_{t+h}}\bm{P}' & \bm{Q}\bm{\hat{\Sigma}_{t+h}}\bm{Q}'
\end{pmatrix}^{-1}
\begin{pmatrix}\tilde{\bm{b}}_{t+h} - \bm{P}\bm{\hat{\mu}}_{t+h}\\ \tilde{\bm{t}}_{t+h}- \bm{Q}\bm{\hat{\mu}}_{t+h}\end{pmatrix} \Big\}.
\end{align*}
This is the joint multivariate Gaussian distribution of $(\tilde{\bm{b}}'_{t+h} ~ \vdots~ \tilde{\bm{t}}'_{t+h})'$. Then from \eqref{4.6} and the properties of the multivariate Gaussian distribution, it follows that
\begin{equation}\label{ex:2.1}
\tilde{\bm{f}}(\tilde{\bm{b}}_{t+h})=\frac{1}{(2\pi)^{\frac{n}{2}}\Big|\bm{P}\bm{\hat{\Sigma}_{t+h}}\bm{P}'\Big|^{\frac{1}{2}}}
\exp \Big\{-\frac{1}{2} (\tilde{\bm{b}}_{t+h} - \bm{P}\bm{\hat{\mu}}_{t+h})' (\bm{P}\bm{\hat{\Sigma}_{t+h}}\bm{P}')^{-1}(\tilde{\bm{b}}_{t+h} - \bm{P}\bm{\hat{\mu}}_{t+h}) \Big\}.
\end{equation}

>>>>

\def\PQ{\begin{pmatrix}\bm{G}\\[-0.2cm]\bm{H}\end{pmatrix}}
\def\bt{\begin{pmatrix}\tilde{\bm{b}}\\[-0.2cm]\tilde{\bm{a}}\end{pmatrix}}

@@

Suppose an unreconciled probabilistic forecast is Gaussian with mean $\hat{\bm \mu}$ and variance-covariance matrix $\hat{\bm \Sigma}$.  The subscripts $t+h|t$ are surpresed for brevity.  The unreconciled density

\begin{equation}
  f(\hat{\bm{y}})=(2\pi)^{-n/2}|\hat{\bm{\Sigma}}|^{-1/2}\exp\left\{-\frac{1}{2}\left[(\hat{\bm y}-\hat{\bm \mu})'\hat{\bm{\Sigma}}^{-1}(\hat{\bm y}-\hat{\bm \mu})\right]\right\}
\end{equation}

After a change in basis

\begin{equation}
f(\tilde{\bm b},\tilde{\bm a})=(2\pi)^{-\frac{n}{2}}\Big|\bm{\hat{\Sigma}}_{t+h}\Big|^{-\frac{1}{2}}\Big|(\bm{S} ~ \vdots~ \bm{R})\Big|\exp\left\{-\frac{1}{2}q\right\}\,,
\end{equation}
where
\begin{equation}
q=(\bm{S}\tilde{\bm{b}}+\bm{R}\tilde{\bm{a}}-\bm{\hat{\mu}})' \bm{\hat{\Sigma}}^{-1}(\bm{S}\tilde{\bm{b}}+\bm{R}\tilde{\bm{a}}-\bm{\hat{\mu}})
\end{equation}

The quadratic form $q$ can be rearranged as
\begin{align*}
q& = 
\Big((\bm{S} ~ \vdots~ \bm{R})\bt-\bm{\hat{\mu}}\Big)' \bm{\hat{\Sigma}}^{-1}\Big((\bm{S} ~ \vdots~ \bm{R})\bt-\bm{\hat{\mu}}\Big),\\
& =
\Big(\bt-(\bm{S} ~ \vdots~ \bm{R})^{-1}\bm{\hat{\mu}}_{t+h}\Big)' \Big[(\bm{S} ~ \vdots~ \bm{R})\bm{\hat{\Sigma}_{t+h}}(\bm{S} ~ \vdots~ \bm{R})'\Big]^{-1}
\Big(\bt-(\bm{S} ~ \vdots~ \bm{R})^{-1}\bm{\hat{\mu}}_{t+h}\Big)\,.
\end{align*}

Recall that
$$
(\bm{S} ~ \vdots~ \bm{R})^{-1} =
\begin{pmatrix}(\bm{R}'_\bot \bm{S})^{-1}\bm{R}'_\bot \\ \cdots \\ (\bm{S}'_\bot \bm{R})^{-1}\bm{S}'_\bot \end{pmatrix} =
\begin{pmatrix}
\bm{G} \\\bm{H}
\end{pmatrix},
$$
where $\bm{G}=(\bm{R}'_\bot \bm{S})^{-1}\bm{R}'_\bot$ and $\bm{H}=(\bm{S}'_\bot \bm{R})^{-1}\bm{S}'_\bot$. Then $q$ can be simplified further as
\begin{align*}
q& =%\frac{1}{(2\pi)^{\frac{n}{2}}\Big|\bm{\hat{\Sigma}}_{t+h}\Big|^{\frac{1}{2}}\Big|\PQ \Big|}
%\exp \Big\{-\frac{1}{2}
\Big[\bt-\PQ\bm{\hat{\mu}}_{t+h}\Big]'%\\[-0.5cm]
\Big[\PQ\bm{\hat{\Sigma}_{t+h}}\PQ'\Big]^{-1}\Big[\bt-\PQ\bm{\hat{\mu}}_{t+h}\Big] %\Big\},
\\[0.5cm]
 & =%\frac{1}{(2\pi)^{\frac{n}{2}}\Big|\PQ\bm{\hat{\Sigma}_{t+h}}\PQ'\Big|^{\frac{1}{2}}}
%\exp \Big\{-\frac{1}{2} 
\begin{pmatrix}\tilde{\bm{b}} - \bm{G}\bm{\hat{\mu}}\\ \tilde{\bm{a}}- \bm{H}\bm{\hat{\mu}}\end{pmatrix}' %&\\[-0.5cm]
%& \hspace*{7cm}
 \Big[\PQ\bm{\hat{\Sigma}_{t+h}}\PQ'\Big]^{-1}\begin{pmatrix}\tilde{\bm{b}} - \bm{G}\bm{\hat{\mu}}\\ \tilde{\bm{a}}- \bm{H}\bm{\hat{\mu}}\end{pmatrix} %\Big\}.
\end{align*}
Since $\Big[\PQ\bm{\hat{\Sigma}_{t+h}}\PQ'\Big] = \begin{pmatrix}
\bm{P}\bm{\hat{\Sigma}_{t+h}}\bm{P}' & \bm{P}\bm{\hat{\Sigma}_{t+h}}\bm{Q}' \\
\bm{Q}\bm{\hat{\Sigma}_{t+h}}\bm{P}' & \bm{Q}\bm{\hat{\Sigma}_{t+h}}\bm{Q}'
\end{pmatrix}$ we have
\begin{align*}
\bm{f_B}(\cdot)&=\frac{1}{(2\pi)^{\frac{n}{2}}\Big|\begin{pmatrix}
  \bm{P}\bm{\hat{\Sigma}_{t+h}}\bm{P}' & \bm{P}\bm{\hat{\Sigma}_{t+h}}\bm{Q}' \\
  \bm{Q}\bm{\hat{\Sigma}_{t+h}}\bm{P}' & \bm{Q}\bm{\hat{\Sigma}_{t+h}}\bm{Q}'
  \end{pmatrix}\Big|^{\frac{1}{2}}}
\exp \Big\{-\frac{1}{2} \begin{pmatrix}\tilde{\bm{b}}_{t+h} - \bm{P}\bm{\hat{\mu}}_{t+h}\\ \tilde{\bm{t}}_{t+h}- \bm{Q}\bm{\hat{\mu}}_{t+h}\end{pmatrix}'\\
&\hspace*{7cm}\begin{pmatrix}
\bm{P}\bm{\hat{\Sigma}_{t+h}}\bm{P}' & \bm{P}\bm{\hat{\Sigma}_{t+h}}\bm{Q}' \\
\bm{Q}\bm{\hat{\Sigma}_{t+h}}\bm{P}' & \bm{Q}\bm{\hat{\Sigma}_{t+h}}\bm{Q}'
\end{pmatrix}^{-1}
\begin{pmatrix}\tilde{\bm{b}}_{t+h} - \bm{P}\bm{\hat{\mu}}_{t+h}\\ \tilde{\bm{t}}_{t+h}- \bm{Q}\bm{\hat{\mu}}_{t+h}\end{pmatrix} \Big\}.
\end{align*}
This is the joint multivariate Gaussian distribution of $(\tilde{\bm{b}}'_{t+h} ~ \vdots~ \tilde{\bm{t}}'_{t+h})'$. Then from \eqref{4.6} and the properties of the multivariate Gaussian distribution, it follows that
\begin{equation}\label{ex:2.1}
\tilde{\bm{f}}(\tilde{\bm{b}}_{t+h})=\frac{1}{(2\pi)^{\frac{n}{2}}\Big|\bm{P}\bm{\hat{\Sigma}_{t+h}}\bm{P}'\Big|^{\frac{1}{2}}}
\exp \Big\{-\frac{1}{2} (\tilde{\bm{b}}_{t+h} - \bm{P}\bm{\hat{\mu}}_{t+h})' (\bm{P}\bm{\hat{\Sigma}_{t+h}}\bm{P}')^{-1}(\tilde{\bm{b}}_{t+h} - \bm{P}\bm{\hat{\mu}}_{t+h}) \Big\}.
\end{equation}
